Think deep for reef resilience

### 2025-07-11 paper/paper_draft.qmd
Auto-format with line breaks for clarity

---
title: Think deep for reef resilience
author:
  - name: Benjamin Grier
    email: grierben777@gmail.com
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          state: QLD
    attributes:
        corresponding: true
  - name: Takuya Iwanaga
    email: t.iwanaga@aims.gov.au
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          city: Cape Cleveland
          state: QLD
    attributes:
        corresponding: true
abstract: |
    Coral reefs globally face unprecedented threats from climate change, with the Great Barrier Reef (GBR) experiencing severe declines in coral cover from thermal stress events. Understanding which reefs exhibit resilience to climate impacts is critical for conservation planning and intervention strategies. A scenario-based population modelling approach was adopted with larval connectivity dynamics, thermal adaptation, and environmental factors to assess coral reef resilience across GBR locations under future climate projections. Using five Global Climate Models (GCMs) spanning five Equilibrium Climate Sensitivities, three distinct reef trajectories are identified: resilient reefs maintaining coral cover through 2050, intermediate reefs with moderate declines, and vulnerable reefs experiencing severe degradation. Contrary to expectations, larval connectivity patterns did not substantially differentiate resilient from vulnerable reefs. Instead, reef depth emerged as the primary predictor of resilience, with deeper reefs benefiting from thermal refugia that reduced bleaching impacts. Initial coral cover also influenced trajectories, with reefs having lower initial cover showing greater relative resilience. Our findings suggest that depth provides critical protection against thermal stress, contrasting current restoration practices that primarily target easily accessible shallow sites. These results have important implications for reef management, indicating that depth refugia should be prioritized for further consideration in conservation planning alongside connectivity considerations to maximize coral survival under continuing climate change.
keywords:
  - Coral reef
  - Climate change
  - Ecosystem resilience
  - Reef depth
  - Reef connectivity
date: last-modified
format:
  elsevier-docx:
    journal:
      name: Journal Name
      formatting: preprint
      cite-style: authoryear
bibliography: references.bib
engine: julia
crossref:
    custom:
    - kind: float
      reference-prefix: SFigure
      key: sfig
---

@@

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all considered reefs (points). The green polygon represents the Far Northern management area, the yellow polygon represents the Cairns/Cooktown management area, the pink polygon represents the Townsville/Whitsunday management area and the blue polygon represents the Mackay/Capricorn management area. The centroid of each reef is represented as a point for visual clarity, and is coloured according to its GBRMPA reefal bioregion. An annotation represents the approximate location of the Australian Institute of Marine Science Townsville site at Cape Cleveland.](/../outputs/ADRIA_results/HighResCoralStress/figs/region_map.png){#fig-region-map width=`{julia} map_width`}

![Equilibrium Climate Sensitivity (ECS) values in °C for the GCMs selected (point values). The pink - dashed box represents the high-certainty ‘very likely’ range for ECS values and the solid red box represents low-certainty ‘likely’ range for ECS values (Forster et al., 2021).](/../outputs/ADRIA_results/HighResCoralStress/figs/ecs_plot.png){#fig-ecs width=`{julia} ecs_width`}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to 2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area. Bracketed x-axis values represent the Spearman's correlation between reef depth and the number of years cover exceeds the carbonate budget threshold. Point colour represents the median reef depth level in metres. Displays results from median timeseries using ACCESS-ESM1-5.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/depth_carbonate_scatter.png){#fig-access-esm1-5-depth-carbonate width=`{julia} carb_width`}

@@

![Reef area-relative coral cover timeseries clustered within each management area, for each GCM analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5 to 97.5 percentile confidence intervals. Each management area is represented five times (for each GCM), with each GCM displayed horizontally.](/../outputs/ADRIA_results/HighResCoralStress/figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories width=`{julia} timeseries_width`}

@@

![Median depth for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each cluster. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/depth_bioregion_violin.png){#fig-access-esm1-5-bioregion-depth width=`{julia} violin_width`}

@@

![Reef Degree Heating Week timeseries clustered within each management area, for each GCM analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5 to 97.5 percentile confidence intervals. Each management area is represented five times (for each GCM), with each GCM displayed horizontally.](/../outputs/ADRIA_results/HighResCoralStress/figs/GCM_dhw_timeseries_plot.png){#sfig-gcm-dhw-trajectories width=`{julia} timeseries_width`}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to 2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area. Bracketed x-axis values represent the Spearman's correlation between reef log (natural) total connectivity strength and the number of years cover exceeds the carbonate budget threshold. Point colour represents log (natural) scaled reef total connectivity strength. Displays results from median timeseries using ACCESS-ESM1-5.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/log_total_strength_carbonate_scatter.png){#sfig-access-esm1-5-conn-carbonate width=`{julia} carb_width`}

@@

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5 to 97.5 percentile confidence intervals. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/bioregion_cover_timeseries.png){#sfig-access-esm1-5-bioregion-timeseries width=`{julia} timeseries_width`}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median cover – DHW correlation level for each cluster. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwcor width=`{julia} violin_width`}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median DHW level for each cluster. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwmean width=`{julia} violin_width`} <!-- ![Mean timeseries DHW tolerance value for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median DHW tolerance value for each cluster. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwtol width=`{julia} violin_width`} --\>

![Initial proportion of coral habitable area occupied by corals for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median proportion of habitable area occupied for each cluster. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/initial_proportion_bioregion_violin.png){#sfig-access-esm1-5-bioregion-initialproportion width=`{julia} violin_width`}

![Log (natural) source – to – sink connectivity ratios for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines within each violin represent the median log (natural) source – to – sink ratio for each cluster and grey points represent the ratio value for each reef. Vertical grey lines represent a log (natural) source – to – sink ratio of 0 indicating equal outgoing and incoming connectivity levels. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/so_to_si_bioregion_violin.png){#sfig-access-esm1-5-bioregion-logsosi width=`{julia} violin_width`}

![Log (natural) total connectivity strength for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines within each violin represent the median log (natural) total connectivity strength for each cluster. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/log_total_strength_bioregion_violin.png){#sfig-access-esm1-5-bioregion-totalconnectivity width=`{julia} violin_width`}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to 2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area. Bracketed x-axis values represent the Spearman's correlation between reef depth and the number of years cover exceeds the carbonate budget threshold. Point colour represents the median reef depth level in metres. Displays results from median timeseries using ACCESS-CM2.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/depth_carbonate_scatter.png){#sfig-access-cm2-depth-carbonate width=`{julia} carb_width`}

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to 2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area. Bracketed x-axis values represent the Spearman's correlation between reef log (natural) total connectivity strength and the number of years cover exceeds the carbonate budget threshold. Point colour represents log (natural) scaled reef total connectivity strength. Displays results from median timeseries using ACCESS-CM2.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/log_total_strength_carbonate_scatter.png){#sfig-access-cm2-conn-carbonate width=`{julia} carb_width`}

@@

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5 to 97.5 percentile confidence intervals. Displays results for ACCESS-CM2 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/bioregion_cover_timeseries.png){#sfig-access-cm2-bioregion-timeseries width=`{julia} timeseries_width`}

![Median depth for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each cluster. Displays results for ACCESS-CM2 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/depth_bioregion_violin.png){#sfig-access-cm2-bioregion-depth width=`{julia} violin_width`}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median DHW level for each cluster. Displays results for ACCESS-CM2 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-cm2-bioregion-dhwmean width=`{julia} violin_width`}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median cover – DHW correlation level for each cluster. Displays results for ACCESS-CM2 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-access-cm2-bioregion-dhwcor width=`{julia} violin_width`}

@@

![Initial proportion of coral habitable area occupied by corals for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median proportion of habitable area occupied for each cluster. Displays results for ACCESS-CM2 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/initial_proportion_bioregion_violin.png){#sfig-access-cm2-bioregion-initialproportion width=`{julia} violin_width`}

>>>>

---
title: Think deep for reef resilience
author:
  - name: Benjamin Grier
    email: grierben777@gmail.com
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          state: QLD
    attributes:
        corresponding: true
  - name: Takuya Iwanaga
    email: t.iwanaga@aims.gov.au
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          city: Cape Cleveland
          state: QLD
    attributes:
        corresponding: true
abstract: |
    Coral reefs globally face unprecedented threats from climate change, with the Great Barrier Reef (GBR) experiencing severe declines in coral cover from thermal stress events. Understanding which reefs exhibit resilience to climate impacts is critical for conservation planning and intervention strategies. A scenario-based population modelling approach was adopted with larval connectivity dynamics, thermal adaptation, and environmental factors to assess coral reef resilience across GBR locations under future climate projections. Using five Global Climate Models (GCMs) spanning five Equilibrium Climate Sensitivities, three distinct reef trajectories are identified: resilient reefs maintaining coral cover through 2050, intermediate reefs with moderate declines, and vulnerable reefs experiencing severe degradation. Contrary to expectations, larval connectivity patterns did not substantially differentiate resilient from vulnerable reefs. Instead, reef depth emerged as the primary predictor of resilience, with deeper reefs benefiting from thermal refugia that reduced bleaching impacts. Initial coral cover also influenced trajectories, with reefs having lower initial cover showing greater relative resilience. Our findings suggest that depth provides critical protection against thermal stress, contrasting current restoration practices that primarily target easily accessible shallow sites. These results have important implications for reef management, indicating that depth refugia should be prioritized for further consideration in conservation planning alongside connectivity considerations to maximize coral survival under continuing climate change.
keywords:
  - Coral reef
  - Climate change
  - Ecosystem resilience
  - Reef depth
  - Reef connectivity
date: last-modified
format:
  elsevier-docx:
    journal:
      name: Journal Name
      formatting: preprint
      cite-style: authoryear
bibliography: references.bib
engine: julia
crossref:
  custom:
  - kind: float
    reference-prefix: SFigure
    key: sfig
editor:
  markdown:
    wrap: 92
---

@@

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points). The green polygon represents the Far Northern management area,
the yellow polygon represents the Cairns/Cooktown management area, the pink polygon
represents the Townsville/Whitsunday management area and the blue polygon represents the
Mackay/Capricorn management area. The centroid of each reef is represented as a point for
visual clarity, and is coloured according to its GBRMPA reefal bioregion. An annotation
represents the approximate location of the Australian Institute of Marine Science Townsville
site at Cape
Cleveland.](/../outputs/ADRIA_results/HighResCoralStress/figs/region_map.png){#fig-region-map
width="`{julia"} map_width\`}

![Equilibrium Climate Sensitivity (ECS) values in °C for the GCMs selected (point values).
The pink - dashed box represents the high-certainty ‘very likely’ range for ECS values and
the solid red box represents low-certainty ‘likely’ range for ECS values (Forster et al.,
2021).](/../outputs/ADRIA_results/HighResCoralStress/figs/ecs_plot.png){#fig-ecs
width="`{julia"} ecs_width\`}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
ACCESS-ESM1-5.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/depth_carbonate_scatter.png){#fig-access-esm1-5-depth-carbonate
width="`{julia"} carb_width\`}

@@

![Reef area-relative coral cover timeseries clustered within each management area, for each
GCM analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2040 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](/../outputs/ADRIA_results/HighResCoralStress/figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
width="`{julia"} timeseries_width\`}

@@

![Median depth for reefs within each bioregion, separated into clusters assigned by
timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median
area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange)
are clusters with intermediate median area-relative cover between 2030 and 2040 within each
bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover
between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each
cluster. Displays results for ACCESS-ESM1-5
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/depth_bioregion_violin.png){#fig-access-esm1-5-bioregion-depth
width="`{julia"} violin_width\`}

@@

![Reef Degree Heating Week timeseries clustered within each management area, for each GCM
analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2040 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](/../outputs/ADRIA_results/HighResCoralStress/figs/GCM_dhw_timeseries_plot.png){#sfig-gcm-dhw-trajectories
width="`{julia"} timeseries_width\`}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef log (natural)
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Point colour represents log (natural) scaled reef total connectivity strength.
Displays results from median timeseries using
ACCESS-ESM1-5.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/log_total_strength_carbonate_scatter.png){#sfig-access-esm1-5-conn-carbonate
width="`{julia"} carb_width\`}

@@

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2040 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2040 within each
bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5
to 97.5 percentile confidence intervals. Displays results for ACCESS-ESM1-5
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/bioregion_cover_timeseries.png){#sfig-access-esm1-5-bioregion-timeseries
width="`{julia"} timeseries_width\`}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster. Displays results
for ACCESS-ESM1-5
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwcor
width="`{julia"} violin_width\`}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters
assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2040
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW level for each cluster. Displays results for ACCESS-ESM1-5
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwmean
width="`{julia"}
violin_width`} <!-- ![Mean timeseries DHW tolerance value for reefs within each bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the median DHW tolerance value for each cluster. Displays results for ACCESS-ESM1-5 GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwtol width=`{julia}
violin_width\`} --\>

![Initial proportion of coral habitable area occupied by corals for reefs within each
bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green)
are clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median proportion of habitable area occupied for each cluster. Displays
results for ACCESS-ESM1-5
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/initial_proportion_bioregion_violin.png){#sfig-access-esm1-5-bioregion-initialproportion
width="`{julia"} violin_width\`}

![Log (natural) source – to – sink connectivity ratios for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median log (natural) source – to – sink ratio for
each cluster and grey points represent the ratio value for each reef. Vertical grey lines
represent a log (natural) source – to – sink ratio of 0 indicating equal outgoing and
incoming connectivity levels. Displays results for ACCESS-ESM1-5
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/so_to_si_bioregion_violin.png){#sfig-access-esm1-5-bioregion-logsosi
width="`{julia"} violin_width\`}

![Log (natural) total connectivity strength for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines within each
violin represent the median log (natural) total connectivity strength for each cluster.
Displays results for ACCESS-ESM1-5
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/bioregion/log_total_strength_bioregion_violin.png){#sfig-access-esm1-5-bioregion-totalconnectivity
width="`{julia"} violin_width\`}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
ACCESS-CM2.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/depth_carbonate_scatter.png){#sfig-access-cm2-depth-carbonate
width="`{julia"} carb_width\`}

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef log (natural)
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Point colour represents log (natural) scaled reef total connectivity strength.
Displays results from median timeseries using
ACCESS-CM2.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/log_total_strength_carbonate_scatter.png){#sfig-access-cm2-conn-carbonate
width="`{julia"} carb_width\`}

@@

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2040 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2040 within each
bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5
to 97.5 percentile confidence intervals. Displays results for ACCESS-CM2
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/bioregion_cover_timeseries.png){#sfig-access-cm2-bioregion-timeseries
width="`{julia"} timeseries_width\`}

![Median depth for reefs within each bioregion, separated into clusters assigned by
timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median
area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange)
are clusters with intermediate median area-relative cover between 2030 and 2040 within each
bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover
between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each
cluster. Displays results for ACCESS-CM2
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/depth_bioregion_violin.png){#sfig-access-cm2-bioregion-depth
width="`{julia"} violin_width\`}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters
assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2040
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW level for each cluster. Displays results for ACCESS-CM2
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-cm2-bioregion-dhwmean
width="`{julia"} violin_width\`}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster. Displays results
for ACCESS-CM2
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-access-cm2-bioregion-dhwcor
width="`{julia"} violin_width\`}

@@

![Initial proportion of coral habitable area occupied by corals for reefs within each
bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green)
are clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median proportion of habitable area occupied for each cluster. Displays
results for ACCESS-CM2
GCM.](/../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/initial_proportion_bioregion_violin.png){#sfig-access-cm2-bioregion-initialproportion
width="`{julia"} violin_width\`}
